Statistical Geometry of the turbulent ISM - I. Field lines

### 2015-01-08 paper.tex
Added some references and motivation to the intro.

\documentclass[useAMS,usenatbib]{mn2e}
%FAG added header file to contain macros
\include{header}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{mathtools}
\usepackage{graphicx,subfigure}
\usepackage{caption}
\usepackage{varioref}
\usepackage{float}
\restylefloat{table}
% If your system does not have the AMS fonts version 2.0 installed, then
% remove the useAMS option.
%
% useAMS allows you to obtain upright Greek characters.
% e.g. \umu, \upi etc.  See the section on "Upright Greek characters" in
% this guide for further information.
%
% If you are using AMS 2.0 fonts, bold math letters/symbols are available
% at a larger range of sizes for NFSS release 1 and 2 (using \boldmath or
% preferably \bmath).
%
% The usenatbib command allows the use of Patrick Daly's natbib.sty for
% cross-referencing.
%
% If you wish to typeset the paper in Times font (if you do not have the
% PostScript Type 1 Computer Modern fonts you will need to do this to get
% smoother fonts in a PDF file) then uncomment the next line
% \usepackage{Times}

@@

\title[Statistical Geometry of the turbulent multi-phase ISM - I. Field lines]{Statistical Geometry of the turbulent ISM - I. Field lines}
\author[C. C.~Evirgen, A.~Shukurov, F.A. Gent and A.~Fletcher]{C.C.~Evirgen$^{1}$\thanks{E-mail:
c.c.evirgen@newcastle.ac.uk (CCE); anvar.shukurov@newcastle.ac.uk (AS); (Fred's email here) andrew.fletcher@newcastle.ac.uk (AF)}, A.~Shukurov$^{1}$, F.A.~Gent$^{2}$ and A.~Fletcher$^{1}$.\\%$^{2}$\footnotemark[1]\thanks{This file has been amended to
%highlight the proper use of \LaTeXe\ code with the class file.
%These changes are for illustrative purposes and do not reflect the
%original paper by A. V. Raveendran.}\\
$^{1}$School of Mathematics and Statistics, Newcastle University,
Newcastle upon Tyne, UK, NE1 7RU\\$^{2}$Fred's address}
\begin{document}
\newcommand{\bvec}[1]{\boldsymbol{#1}}
\newcommand{\avg}[1]{\left<\bvec{#1}\right>_{l}}
\date{Accepted -. Received -; in original form -}

@@

\section{Introduction}
The results presented in this paper are based on simulations of the multiphase interstellar medium (ISM) by Gent \textsl{et al} in \citep{gent1} and \citep{gent2}, %further referencing to Paper I & II needed here.
focusing on the the B1$\Omega$s model of the ISM. This model features supernova(SN)-driven turbulence with gravity, differential rotation, SN explosions (SNe), radiative cooling, photoelectric heating and other transport processes, where the dynamo is in a kinematic statistically-steady state \citep{gent1,gent2}.\\
A snapshot of the ISM at approximately $t=1.4$ Gyr is used. Its characteristics, in terms of the fractional volumes of the phases, is representative of the steady-state phase of the dynamo. In addition, there are two potential SN remnants, which provide isolated structures of hot gas. This enhances the illustration of the effect that hot gas has on the mean magnetic field. \\
The multi-phase structure is defined in terms of temperature and density. However, an equivalent definition is given in terms of entropy, which has an explicit defintion linking density and temperature \citep{gentphd}. The cold, warm and hot phases are defined by the entropy ranges, $s\leq3.7\cdot10^8$, $3.7\cdot10^8<s<23.2\cdot10^8$ and $s>23.2\cdot10^8$ erg g$^{-1}$ K$^{-1}$, respectively. 

@@

When considering the mean-field decomposition of the magnetic field, the notation used by \citep{gent2} is adopted.

@@

where $l\approx50$pc is half the integral scale of random motions \citep{gent2}. This value is adopted for this paper based on the justification given in \citep{gent2}.  Preliminary analysis does not show significant sensitivity of the mean or random field to variations in $l$ within the range $0<l<100$pc. However, future work will consider analysis of isolated structures in the ISM. A more detailed analysis of length and correlation scales will be carried out. However, the current definition, $l=50$pc, is sufficient for the purposes of this paper. 

@@

Field lines of the magnetic field give a qualitative description of the magnetic field in the ISM. We introduce sampling of physical variables along the field lines of the magnetic field, to obtain a quantitative description of the observables. Observables, such as entropy or pulsar RM measures \citep{st02} are frequently used to derive other observables or statistics.  The use of sampling along field lines will enable the analysis of observables across many snapshots, whilst avoiding the use of time averages, which is not sensible for observables associated to a turbulent flow \citep{tnl}. For example, the wavelet transform methods used by Stepanov \textsl{et al} \citep{st02} are used to derive the Galactic magnetic field from pulsar RM data. Field lines of the Galactic magnetic field can be derived from this analysis and it can be further analysed using sampling along the field lines.\\

@@

Even though the visualisation is informative, it is necessary to determine whether the field lines are sensitive to the phases of the ISM. Consequently, we compare the characteristics of the field lines against straight lines launched in the positive $y$-direction, from the seed points used to compute the field lines. This test is devised using ideas from the field of stereology \citep{stereo}, lineal analysis in particular. The straight lines used in the test case will have the same general characteristics as the field lines, since they both traverse the computational domain in the positive $y$-direction. This ensures that the lines are representative of the field lines being sampled. Despite having the same general characteristics, the straight lines should not be sensitive to the phases of the ISM; they will not prefer to stay in a particular phase, whilst avoiding others. Therefore, comparison of the PDFs of the observables sampled along the field lines and the straight lines, should identify sensitivity (or lack thereof) of the field lines to the phases of the ISM.\\

@@

CCE thanks Dr Frederick Gent for helpful discussions of his work.

\begin{thebibliography}{99}
\bibitem[\protect\citeauthoryear{Gent \textsl{et al}}{2013}]{gent1} Gent F.A., Shukurov A., Sarson G.R., Fletcher A., Mantere M.J., 2012, MNRAS, 432, 1396-1423
\bibitem[\protect\citeauthoryear{Gent \textsl{et al}}{2012}]{gent2} Gent F.A., Shukurov A., Sarson G.R., Fletcher A., Mantere M.J., 2012, MNRASL
\bibitem[\protect\citeauthoryear{Tennekes and Lumley}{1972}]{tnl} Tennekes H.,~Lumley J.L.,\emph{A First Course in Turbulence},~1972,~MIT Press
\bibitem[\protect\citeauthoryear{Gent}{2012}]{gentphd} Gent F.A.,~PhD Thesis:~\emph{Supernova Driven Turbulence in the Interstellar Medium} ,~2012,~Newcastle University School of Mathematics and Statistics%,\url{<http://hdl.handle.net/10443/1755
%<http://hdl.handle.net/10443/1755>}
\bibitem[\protect\citeauthoryear{Stepanov \textsl{et al}}{2002}]{st02} Stepanov R.,~Frick P.,~Shukurov A.,~Sokoloff D.,~\emph{Wavelet tomography of the Galactic magnetic field},~2002,~A\&A,~391,~361-368
\bibitem[\protect\citeauthoryear{Baddeley and Jensen}{2005}]{stereo} Baddeley A.,~Jensen E.B.,~\emph{Stereology for Statisticians},~2005,~Chapman \& Hall/CRC
\end{thebibliography}

>>>>

\documentclass[useAMS,usenatbib]{mn2e}
%FAG added header file to contain macros
\include{header}
%\usepackage{amsmath}
%\usepackage{bm}
%\usepackage{mathtools}
%\usepackage{graphicx,subfigure}
%\usepackage{caption}
%\usepackage{varioref}
%\usepackage{float}
%\restylefloat{table}
% If your system does not have the AMS fonts version 2.0 installed, then
% remove the useAMS option.
%
% useAMS allows you to obtain upright Greek characters.
% e.g. \umu, \upi etc.  See the section on "Upright Greek characters" in
% this guide for further information.
%
% If you are using AMS 2.0 fonts, bold math letters/symbols are available
% at a larger range of sizes for NFSS release 1 and 2 (using \boldmath or
% preferably \bmath).
%
% The usenatbib command allows the use of Patrick Daly's natbib.sty for
% cross-referencing.
%
% If you wish to typeset the paper in Times font (if you do not have the
% PostScript Type 1 Computer Modern fonts you will need to do this to get
% smoother fonts in a PDF file) then uncomment the next line
% \usepackage{Times}

@@

\title[Statistical Geometry of the turbulent multi-phase ISM - I. Field lines]{Statistical Geometry of the turbulent ISM - I. Field lines}
\author[C. C.~Evirgen, A.~Shukurov, F.A. Gent and A.~Fletcher]{C.C.~Evirgen$^{1}$\thanks{E-mail:
c.c.evirgen@newcastle.ac.uk (CCE); anvar.shukurov@newcastle.ac.uk (AS); 
f.gent@shef.ac.uk (FAG); andrew.fletcher@newcastle.ac.uk (AF)}, A.~Shukurov$^{1}$, F.A.~Gent$^{2}$ and A.~Fletcher$^{1}$.\\%$^{2}$\footnotemark[1]\thanks{This file has been amended to
%highlight the proper use of \LaTeXe\ code with the class file.
%These changes are for illustrative purposes and do not reflect the
%original paper by A. V. Raveendran.}\\
$^{1}$School of Mathematics and Statistics, Newcastle University,
Newcastle upon Tyne, UK, NE1 7RU\\
$^{2}$SP$^{\,2}$\!RC, School of Mathematics and Statistics, University of Sheffield, 
Sheffield, UK, S3 7RH}
\begin{document}
\newcommand{\bvec}[1]{\boldsymbol{#1}}
\newcommand{\avg}[1]{\left<\bvec{#1}\right>_{l}}
\date{Accepted -. Received -; in original form -}

@@

\section{Introduction}
%FAG rephrase
%The results presented in this paper are based on simulations of the multiphase interstellar medium (ISM) by Gent \textsl{et al} in \citep{GSFSM13} and \citep{GSSFM13}, %further referencing to Paper I & II needed here.
The results presented in this paper are based on simulations of the multiphase 
interstellar medium (ISM) reported in \citet{GSFSM13} and \citet{GSSFM13}, %further referencing to Paper I & II needed here.
subsequently referred to as \HD\, and \MHD, respectively.
%end FAG
%FAG rephrase
In particular we consider a model, identified as B1$\Omega$s in \MHD.
Supernova(SN)-driven turbulence is applied to ISM in a section of a spiral
galaxy differentially rotating, stratified about the galactic disk and subject
to, radiative cooling, photoelectric heating and other transport processes.
The rate and distribution of SN explosions, gravity and angular momentum are 
consistent with observational parameters for the solar neighbourhood.
A nano Gauss seed magnetic field is applied and amplified by dynamo until it
saturates with a mean magnetic stength of a few micro Gauss, remarkably 
consistent with observed estimates for the Milky Way.

Models of the magnetized ISM \citep{Heitsch01,M-LBKA05} 
%FAG maybe also include Henley et al 2015 end FAG
have explored the 
effects of magnetic field on the localized ISM, including the fluctuation 
dynamo, but without the effects of stratification and the large scale 
structure induced by the differential rotation of the disk.
Large scale structure in the stratified ISM has been investigated by
\citet{AB05a}, but without differential rotation the dynamo to grow the field
to full strength is absent, so an imposed field is applied. 
\citet{Hanasz05,Dobbs08} and \citet{Hanasz09} apply magnetic fields to global 
disk models, including a dynamo, but due to numerical constraints must limit or 
exclude vertical stratification and multi-phase gas composition.
\citet{Korpi99a} and \cite{Gressel08} conducted earlier simulations with the
same approach as described here, but with lower resolution and higher 
diffusion were unable to follow the dynamo through to saturation.
The advantage of the results from \MHD\ is that the magnetic field produced
is not imposed, but evolves dynamically under realistic physical processes
governing the formation of the thermo-hydrodynamic structure of the ISM.
From this point of view it is of great value for appreciating the 
multi-phase structure of the magnetized ISM and for comparison with observation.

%focusing on the the B1$\Omega$s model of the ISM. This model features supernova(SN)-driven turbulence with gravity, differential rotation, SN explosions (SNe), radiative cooling, photoelectric heating and other transport processes, where the dynamo is in a kinematic statistically-steady state \citep{GSFSM13,GSSFM13}.\\
%end FAG
A snapshot of the ISM at approximately $t=1.4$ Gyr is used. Its characteristics, in terms of the fractional volumes of the phases, is representative of the steady-state phase of the dynamo. In addition, there are two potential SN remnants, which provide isolated structures of hot gas. This enhances the illustration of the effect that hot gas has on the mean magnetic field.

The multi-phase structure is defined in terms of temperature and density. However, an equivalent definition is given in terms of entropy, which has an explicit defintion linking density and temperature \citep{Gent12}. The cold, warm and hot phases are defined by the entropy ranges, $s\leq3.7\cdot10^8$, $3.7\cdot10^8<s<23.2\cdot10^8$ and $s>23.2\cdot10^8$ erg g$^{-1}$ K$^{-1}$, respectively. 

@@

When considering the mean-field decomposition of the magnetic field, the notation used by \citep{GSSFM13} is adopted.

@@

where $l\approx50$pc is half the integral scale of random motions \citep{GSSFM13}. This value is adopted for this paper based on the justification given in \citep{gent2}.  Preliminary analysis does not show significant sensitivity of the mean or random field to variations in $l$ within the range $0<l<100$pc. However, future work will consider analysis of isolated structures in the ISM. A more detailed analysis of length and correlation scales will be carried out. However, the current definition, $l=50$pc, is sufficient for the purposes of this paper. 

@@

Field lines of the magnetic field give a qualitative description of the magnetic field in the ISM. We introduce sampling of physical variables along the field lines of the magnetic field, to obtain a quantitative description of the observables. Observables, such as entropy or pulsar RM measures \citep{SSFS02} are frequently used to derive other observables or statistics.  The use of sampling along field lines will enable the analysis of observables across many snapshots, whilst avoiding the use of time averages, which is not sensible for observables associated to a turbulent flow \citep{TL72}. For example, the wavelet transform methods used by Stepanov \textsl{et al} \citep{st02} are used to derive the Galactic magnetic field from pulsar RM data. Field lines of the Galactic magnetic field can be derived from this analysis and it can be further analysed using sampling along the field lines.\\

@@

Even though the visualisation is informative, it is necessary to determine whether the field lines are sensitive to the phases of the ISM. Consequently, we compare the characteristics of the field lines against straight lines launched in the positive $y$-direction, from the seed points used to compute the field lines. This test is devised using ideas from the field of stereology \citep{BJ04}, lineal analysis in particular. The straight lines used in the test case will have the same general characteristics as the field lines, since they both traverse the computational domain in the positive $y$-direction. This ensures that the lines are representative of the field lines being sampled. Despite having the same general characteristics, the straight lines should not be sensitive to the phases of the ISM; they will not prefer to stay in a particular phase, whilst avoiding others. Therefore, comparison of the PDFs of the observables sampled along the field lines and the straight lines, should identify sensitivity (or lack thereof) of the field lines to the phases of the ISM.\\

@@

%FAG now moot
%CCE thanks Dr Frederick Gent for helpful discussions of his work.
%end FAG

%FAG include bibliography file refs.bib and amended cite codes 
%\begin{thebibliography}{99}
%\bibitem[\protect\citeauthoryear{Gent \textsl{et al}}{2013}]{GSFSM13} Gent F.A., Shukurov A., Sarson G.R., Fletcher A., Mantere M.J., 2012, MNRAS, 432, 1396-1423
%\bibitem[\protect\citeauthoryear{Gent \textsl{et al}}{2012}]{GSSFM13} Gent F.A., Shukurov A., Sarson G.R., Fletcher A., Mantere M.J., 2012, MNRASL
%\bibitem[\protect\citeauthoryear{Tennekes and Lumley}{1972}]{TL72} Tennekes H.,~Lumley J.L.,\emph{A First Course in Turbulence},~1972,~MIT Press
%\bibitem[\protect\citeauthoryear{Gent}{2012}]{Gent12} Gent F.A.,~PhD Thesis:~\emph{Supernova Driven Turbulence in the Interstellar Medium} ,~2012,~Newcastle University School of Mathematics and Statistics%,\url{<http://hdl.handle.net/10443/1755
%<http://hdl.handle.net/10443/1755>}
%\bibitem[\protect\citeauthoryear{Stepanov \textsl{et al}}{2002}]{SSFS02} Stepanov R.,~Frick P.,~Shukurov A.,~Sokoloff D.,~\emph{Wavelet tomography of the Galactic magnetic field},~2002,~A\&A,~391,~361-368
%\bibitem[\protect\citeauthoryear{Baddeley and Jensen}{2005}]{BJ04} Baddeley A.,~Jensen E.B.,~\emph{Stereology for Statisticians},~2005,~Chapman \& Hall/CRC
%\end{thebibliography}
  \bibliographystyle{mn2e}      % basic style, author-year citations
  \bibliography{refs}
% name your BibTeX data base
  \label{lastpage}
%end FAG
